### 2021-09-24 documentos/math.tex
App0
App1

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}
\afnote{Por ahora uso este $\ket{\rho}$}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\ket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \ket{\sigma_{\alpha}}$, the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \dyad{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\ket{\sigma_{\alpha}}$ correspond to the vectorized elements of the Pauli basis and satisfy the usual relation $ \braket{{\sigma_{\alpha}}}{{\sigma_{\alpha'}}} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2 ~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Moreover, observe that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel, with Kraus operators given by $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}. In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see Appendix for details),
%
\begin{equation}
 \mathcal{D} \ket{\sigma_{\alpha}}= \lambda_{\alpha} \ket{\sigma_{\alpha}},
\end{equation}
% 
with eigenvalues given by 
% 
\begin{equation}
 \lambda_{\alpha} = \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
%
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes H$ and $H$ is the Hadamard matrix,
%
\begin{equation}
 H=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}
% 
\afnote{Podemos definir la matriz de Hadamard con un factor $2^{-1/2}$, o sea $H'=H/\sqrt{2}$ para evitar cargar factores extra hasta el final, de esta manera tenemos simplemente $\san=H^{\otimes 2N}$. Qué opinan?}

\subsubsection{$N$ qubits}
% 
Analogously to the single qubit case, one can define the $N$-qubits Pauli map $\mathcal{E}_N$
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}_N[\rho ]=\frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \ket{\rho}=\frac{1}{2^N}\sum_{\vec{\alpha}} r_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
where $\ket{\sigma_{\vec{\alpha}}}=\ket{\sigma_{\alpha_1}}\otimes\dots \otimes\ket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation $\braket{{\sigma_{\vec{\alpha}}}}{{\sigma_{\vec{\alpha}'}}} = \tr \left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N ~\delta_{\vec{\alpha}\vec{\alpha}'}$. The matrix representation of this map is given by
% 
\begin{equation}
\hat{\mathcal{E}}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} \dyad{\sigma_{\vec{\alpha}}}.
% \label{Map_Pauli}
\end{equation}
%
As in the previous case, the Choi matrix $\mathcal{D}_N$ may be written in terms of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}}  \bigotimes_{j=1}^N \sigma_{\alpha_j} \otimes \sigma_{\alpha_j}^*.
%  \label{Choi_1}
\end{equation}
% 
After several steps, it is possible to prove that analogously to the single qubit case, $\mathcal{D}_N$ is diagonal in the $N$-qubits Pauli basis
%
\begin{equation}
 \mathcal{D}_N \ket{\sigma_{\vec{\alpha}}}= \lambda_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
with eigenvalues $\lambda_{\vec{\alpha}}$ given by 
% 
\begin{equation}
 \lambda_{\vec{\alpha}} = \sum_{\vec{\beta}} \san_{\vec{\alpha}\vec{\beta}}\tau_{\vec{\beta}},
\end{equation}
%
where 
% 
\begin{equation}
 \san=\sa^{\otimes N}=\frac{1}{2^N}\left(H\otimes H\right)^{\otimes N}=\left(\frac{H}{\sqrt{2}}\right)^{\otimes 2N}.
\end{equation}
%

@@

\subsubsection{Appendix - Diagonalization of Choi Matrix $\mathcal{D}_N$}

>>>>

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}
\afnote{Por ahora uso este $\kket{\rho}$}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\kket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \kket{\sigma_{\alpha}}$, the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \ddyada{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\kket{\sigma_{\alpha}}$ correspond to the vectorized elements of the Pauli basis and satisfy the usual relation $ \bbrakket{{\sigma_{\alpha}}}{{\sigma_{\alpha'}}} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2 ~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Moreover, observe that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel, with Kraus operators given by $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}. In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see Appendix for details),
%
\begin{equation}
 \mathcal{D} \kket{\sigma_{\alpha}}= \lambda_{\alpha} \kket{\sigma_{\alpha}},
\end{equation}
% 
with eigenvalues given by 
% 
\begin{equation}
 \lambda_{\alpha} = \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
%
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes H$ and $H$ is the Hadamard matrix,
%
\begin{equation}
 H=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}
% 
% 
\afnote{Podemos definir la matriz de Hadamard con un factor $2^{-1/2}$, o sea $H'=H/\sqrt{2}$ para evitar cargar factores extra hasta el final, de esta manera tenemos simplemente $\san=H^{\otimes 2N}$. Qué opinan?}

\subsubsection{$N$ qubits}
% 
Analogously to the single qubit case, one can define the $N$-qubits Pauli map $\mathcal{E}_N$
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}_N[\rho ]=\frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \kket{\rho}=\frac{1}{2^N}\sum_{\vec{\alpha}} r_{\vec{\alpha}} \kket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
where $\kket{\sigma_{\vec{\alpha}}}=\kket{\sigma_{\alpha_1}}\otimes\dots \otimes\kket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation $\bbrakket{{\sigma_{\vec{\alpha}}}}{{\sigma_{\vec{\alpha}'}}} = \tr \left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N ~\delta_{\vec{\alpha}\vec{\alpha}'}$. The matrix representation of this map is given by
% 
\begin{equation}
\hat{\mathcal{E}}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} \ddyada{\sigma_{\vec{\alpha}}}.
% \label{Map_Pauli}
\end{equation}
%
As in the previous case, the Choi matrix $\mathcal{D}_N$ may be written in terms of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}}  \bigotimes_{j=1}^N \sigma_{\alpha_j} \otimes \sigma_{\alpha_j}^*.
%  \label{Choi_1}
\end{equation}
% 
After several steps, it is possible to prove that analogously to the single qubit case, $\mathcal{D}_N$ is diagonal in the $N$-qubits Pauli basis
%
\begin{equation}
 \mathcal{D}_N \kket{\sigma_{\vec{\alpha}}}= \lambda_{\vec{\alpha}} \kket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
with eigenvalues $\lambda_{\vec{\alpha}}$ given by 
% 
\begin{equation}
 \lambda_{\vec{\alpha}} = \sum_{\vec{\beta}} \san_{\vec{\alpha}\vec{\beta}}\tau_{\vec{\beta}},
\end{equation}
%
where 
% 
\begin{equation}
 \san=\sa^{\otimes N}=\frac{1}{2^N}\left(H\otimes H\right)^{\otimes N}=\left(\frac{H}{\sqrt{2}}\right)^{\otimes 2N}.
\end{equation}
%
%

@@

\subsubsection{APPENDIX: Diagonalization of Choi Matrix $\mathcal{D}_N$}
% 
In order to simplify the derivation of the relations, let us employ pairs of binary indices instead of a single quaternary, i.e. $\alpha\to\mu+2\nu$. In this way we can write the elements of the Pauli basis $(\sigma_{0},\sigma_{1},i\sigma_{2},\sigma_{3})$ in the standard and vectorized forms, compactly as
% 
\begin{equation}
\sigma_{\mu\nu}=\sum_{j=0}^{1} (-1)^{j \mu}\dyad{j}{j\oplus \nu},
\end{equation}
% 
and
% 
\begin{equation}
\kket{\sigma_{\mu\nu}}=\sum_{j=0}^{1} (-1)^{j \mu}\ket{j,j\oplus \nu}.
\end{equation}
% 
In addition, recall the inverse relation between the computational and Pauli elements
% 
\begin{equation}
\ket{l,l\oplus n}=\frac{1}{2}\sum_{m=0}^{1} (-1)^{ml}\kket{\sigma_{mn}}.
\label{PauToComp}
\end{equation}
% 
On the other hand, the matrix form of an arbitrary Pauli map $\mathcal{E}$ may be written as
% 
\begin{align}
\hat{\mathcal{E}} & = \frac{1}{2}\sum_{\mu\nu=0}^{1} \tau_{\mu\nu} \ddyada{\sigma_{\mu\nu}},\\
& = \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} \dyad{j,j\oplus\nu}{k,k\oplus\nu}.
% \label{Map_Pauli}
\end{align}
% 
After applying the reshuffling operation on $\hat{\mathcal{E}}$, we obtain the Choi matrix associated to the map. It reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} \dyad{j,k}{j\oplus\nu,k\oplus\nu}.
\label{Choi_a1}
\end{equation}
%

Let us appply the index relabeling $k\to j\oplus m$, then the Choi matrix holds
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu} (-1)^{\mu m} \dyad{j,j\oplus m}{j\oplus\nu,j\oplus\nu\oplus m}.
\label{Choi_a1}
\end{equation}
%
By using the relation between computational and Pauli elements (Eq. \ref{PauToComp}), we have
% 
\begin{align*}
 \mathcal{D} & = \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu}(-1)^{\mu m}  \left( \frac{1}{2}\sum_{k=0}^{1} (-1)^{kj}\kket{\sigma_{km}}\right)\left(\frac{1}{2}\sum_{n=0}^{1} (-1)^{n(j\oplus\nu)}\bbra{\sigma_{nm}}\right),\\
 & = \frac{1}{8}\sum_{jkmn\mu\nu} \tau_{\mu\nu}(-1)^{\mu m+kj+n(j\oplus\nu)}  \ddyad{\sigma_{km}}{\sigma_{nm}},\\
 & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \left(\sum_{j} (-1)^{j(k+n)} \right) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
 & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} (2\delta_{kn}) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
 & = \frac{1}{4}\sum_{mn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \ddyad{\sigma_{nm}}{\sigma_{nm}},\\
 & = \frac{1}{2}\sum_{mn} \left(\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu}\right) \ddyad{\sigma_{nm}}{\sigma_{nm}},
% \label{Choi_a1}
\end{align*}
%
Then $\mathcal{D}$ is diagonal in the Pauli basis, with eigenvalues given by
% 
\begin{equation}
\lambda_{mn}=\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu},
\end{equation}
% 
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes H$ and $H$ is the Hadamard matrix.

\afnote{Queda faltando pasar a limpio la generalizacion para $N$ qubits.}
